Generalised gravitational burst generation with Generative Adversarial Networks

### 2020-12-04 Paper_draft/main.tex
Add a few comments and fix some typos

% What this paper does on GANs in 1 paragraph - the point of the paper
%
In this work we aim to explore the use of \ac{ML} to generate and interpret
unmodeled \ac{GW} burst waveforms. Using the generative machine learning
model, \acp{GAN}, we train on five classes of waveforms in the time domain. Working on the assumption that \acp{GAN} construct smooth
 high dimensional vector spaces between their input and output, we can then
explore the space between the five classes to construct new
hybrid waveforms. As all the computationally expensive
processes occur during training, once trained, the model is able to
generate waveforms in fractions of a second and produce waveforms that are difficult to generate with current
techniques. These new varieties of waveforms can then be used to diagnose
detection algorithms, gain new insight into sources of \ac{GW}
bursts and  allow us to better train our algorithms on a
broader range of possible signals and therefore enhance our detection ability.

% the structure of the paper
%
This paper is organised as follows. In \cref{ML overview} we introduce the basic ideas of machine learning and discuss the choice of algorithm we used. In \cref{Method} we describe the training data and the details of the model. We present the results of the GAN in \cref{results} and show how unmodeled signals can be produced by interpolating and sampling within latent and class spaces. In \cref{cnn classifier} we show that a convolutional neural network classifier can be trained to distinguish between sets of our GAN generated waveforms from noise only cases. We conclude with a summary of the presented work in \cref{conclusions}.

@@

\ac{ML} algorithms aim to learn apparent relationships held within given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the model learning on
past experience to make decisions on future events. Artificial neural networks are universal function approximators that are built from many single
processing units called neurons. The simplest neural network is the perceptron
layer~\cref{fig:perceptron} which shows a single neuron that takes a vector of real
inputs $x_{i},\ldots, x_{n}$
and maps them to an output according to the linear function, 
%
\begin{align}
f(x) = \sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
where $w$ and $b$ are the weights and bias and $\sigma$
denotes the activation function. The weights are numbers which can be thought
of as the strength between connected neurons. The output of a neuron is defined by its activation function which controls how the neuron `fires' depending on its input. Some examples of commonly used activation functions are shown in~\cref{fig:activations}. It is often useful to introduce a bias, $b$, such that the neuron remains inactive
above zero but is active when the sum reaches a defined threshold.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generative Adversarial Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% basic intro to GANs
%
A subset of deep learning that has seen fruitful development in recent years
are generative adversarial networks \acp{GAN}~\cite{Goodfellow2014}. These unsupervised algorithms learn patterns in a
given training data set using an adversarial process. The generations from
\acp{GAN} are currently state-of-the-art in fields such as high quality image
fidelity~\cite{brock2018large,karras2019analyzing}, text-to-image
translation~\cite{reed2016generative}, and video
prediction~\cite{liang2017dual} as well as time series
generations~\cite{esteban2017realvalued}.
%
% Basic components of a GAN
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
network that is set up to distinguish between real and fake data and a
generator network that produces fake versions of the real data. The generator model performs a mapping from a fixed length vector $\mathbf{z}$ to its
representation of the data. The input vector is drawn randomly from a Gaussian distribution which is referred to as a latent space comprised of latent variables. The latent space is a compressed representation of a data distribution which the generator applies meaning to during training. Sampling points from this space allows the generator to produce a variety of different generations, with different points corresponding to different features in the generations. The discriminator maps its input $\mathbf{x}$ to a probability that the input came from either the training (real) data or
generator (fake).
\begin{comment}
Chris' definition of latent space if needed: It maps the intrinsic
variation of the training space onto the Gaussian latent space distribution.
\end{comment}
%
% training a GAN
%
During training, the discriminator and generator are updated using batches of data. Random latent vectors are given to the generator to produce a batch of fake samples and an equal batch of real samples is taken from the training data. The discriminator makes predictions on the real and fake samples and the model is updated through minimising the binary cross-entropy function \cite{Goodfellow-et-al-2016}
%
\begin{equation}
    L = -y \log(\hat{y}) - (1 - y) \log(1-\hat{y}),
    \label{eqn:crossentropy}
\end{equation}
where $\hat{y}$ is the network prediction and $y$ is the true output. While training the discriminator, D, on real data, we set $y = 1$ and $\hat{y} = D(\mathbf{x})$ which from \cref{eqn:crossentropy} gives $L(D(\mathbf{x}),1) = \log(D(\mathbf{x}))$. While training on fake data produced by the generator, G, $y = 0$ and $\hat{y} = D(G(\mathbf{z}))$ and so, $L(D(G(\mathbf{z})),0) = \log(1-(D(G(\mathbf{z}))))$. Since the objective of the discriminator is to correctly classify fake and real data these equations should be maximised, while the goal of the generator should be to minimize these equations. This gives the \ac{GAN} loss as
%
\begin{equation}
   \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{\text{data}}}(\mathbf{x})} [\text{log} D(\mathbf{x})] + \mathbb{E}_{\mathbf{z} \sim p_{\mathbf{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z})))],
 \label{equation:GANloss}
 \end{equation}
where $p_{\text{\text{data}}}(\mathbf{x})$ is the distribution of real data and $p_{\text{z}}(\mathbf{z})$ is the distribution of the latent space.

\subsection{Training stages}
%
Training a GAN is involves updating both the discriminator and generator in stages. First, the discriminator is updated on real instances from the training set. We set the true label $y=1$ and calculate the loss with respect to the predictions $\hat{y}$ via \cref{equation:GANloss}. Stochastic gradient descent is used to minimize the loss which has reduced to $L_D(\textrm{real}) = -\log(\hat{y})$. The discriminator is then trained on fake instances taken from the generator where we set $y=0$ and minimize $L_D(\textrm{fake}) = -\log(1-\hat{y})$. To train the generator, we use a composite model of the generator and discriminator and allow the gradients to flow through this entire model. At this step we set the labels for the generator as real $y=1$ and minimize $L_G = -\log(\hat{y})$. The act of switching the labels and minimizing this form of loss is identical to the previously mentioned generator loss but is beneficial as in practice this function provides more useful gradients. This tweak to the generator loss is called non-saturating generator loss and was reported in the original GAN paper \cite{Goodfellow2014}. It was also shown in that paper that if the generator and discriminator can no longer improve, then the discriminator can no longer distinguish between real and fake i.e. $D(x) = \frac{1}{2}$.

As \acp{GAN} are trained by updating one model at the expense of the other, they can be hard to train. GANs attempting to replicate complicated structures that do not
have the necessary architecture either struggle to produce results at all or
fall into the common failure mode know as mode collapse; where the generator
produces a small variety of samples or simply memorises the training set. The goal of GAN training is to find an equilibrium between the two models, if this cannot be found then it is said that the \ac{GAN} has failed to converge. One way to diagnose \acp{GAN} during the development process is to keep track of the loss and accuracy over time. Loss plots, for example, as seen in \cref{fig:lossplot} can help to identify common failure modes or to check if the \ac{GAN} has indeed converged. Accuracy is another metric that may be used to monitor convergence and is defined as the number of correct predictions made divided by total number of predictions. There is currently no notion of early stopping in \acp{GAN}, instead, training is halted after convergence and by visually inspecting the generations. 
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/losses.png}
    \caption{Plot of the discriminator and generator loss and accuracy as a function of epochs. Early in training the losses oscillates as both models attempt to find their equilibrium, after which, both losses vary around a point which signifies stable training. Accuracy's on the real and fake data are similar showing that neither model is stronger than the other.}
    \label{fig:lossplot}
\end{figure}

@@

% introduce CGANs and ACGANs
%
To gain more control over what a GAN is able to generate, a conditional variant
of \acp{GAN} named \acp{CGAN}~\cite{cgan} was introduced by feeding in extra
information into the generator and discriminator such as a class label or
attribute label, $\mathbf{c}$. This simple addition has shown to work well in practice, for instance in image-to-image translation~\cite{isola2016imagetoimage}. We defined the classes by one-hot encoding, that is, each class resides at the corner points of a 5 dimensional cube. For example $\mathbf{c}=[0,1,0,0,0]$ represents the ringdown signal class. The training data and labels are drawn from a joint distribution $p_{\text{data}}(\mathbf{x},\mathbf{c})$, whereas when generating fake data we sample from $\mathbf{c}$ and $p_{\mathbf{z}}(\mathbf{z})$ independently. \cref{equation:GANloss} is modified to include the class labels 
~
\begin{equation}
   \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{\text{data}}}(\mathbf{x})} [\text{log} D(\mathbf{x|c})] + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z|c})))].
 \label{equation:CGANloss}
 \end{equation}
Fig. \ref{fig:gan_comparison} shows the differences in inputs and outputs of a GAN compared with a \ac{CGAN}. We will be using a conditional GAN for this study.

@@

% introduce the training data
%
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/training-sample.png}
    \caption{Examples of the five different waveforms that were used in training the \ac{GAN} for this study. Values of the parameters were selected randomly from uniform distributions from \cref{Tab:training_parms}}
    \label{fig:training_waveforms}
\end{figure}
% introduce the signal models
%
We propose a signal generation scheme using a \ac{CGAN} trained on burst-like
waveforms. We call this \texttt{McGANn}\footnote{https://github.com/jmcginn/McGANn} and it is a \ac{CGAN} trained on five signal classes each spanning a range
of prior signal parameters. The signal classes are:

@@

The design of the networks is influenced by \cite{Radford2015} in which they use a deep convolution generative adversarial network architecture \ac{DCGAN}. The generator model is fully convolutional,
upsampled using strided transposed convolutions
with batch normalisation in the
first layer and ReLU activations throughout with the exception of the a linear activation for
the output layer. The use of a linear activation guarantees the output can have negative and positive outputs. Each transposed
convolutional layer uses a kernel size of 18 and stride of 2. The discriminator network
mirrors that of the generator without batch normalization, using
LeakyReLU activations, SpatialDropout, and a 2-stride convolution for
downsampling. The discriminator output is a single node with sigmoid activation that can be interpreted as a probability of the the signal being real and both models are trained with binary cross
entropy \cref{eqn:crossentropy}. The full architecture description can be seen in~\cref{Tab:gan_training_parms}.

All models were designed with the Python Keras library \cite{} and TensorFlow \cite{} and ran on a GeForce RTX 2080 Ti GPU. We train the networks for 500 epochs which takes $\mathcal{O}(10)$ hours and save the model at each epoch. We choose an appropriate model by visually inspecting the generations at a point of convergence on the loss plot. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results} \label{results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

While the \ac{GAN} is trained on distinct one-hot encoded classes, we may test arbitrary points in the 5-dimensional class space to produce indistinct or hybrid waveforms. In order to explore the class-space, in \cref{fig:c_interp} we show results where the latent vector is held constant but we instead linearly
interpolate within the one-hot encoded class-space between pairs of the well-defined training class locations. In this scenario we highlight that the \ac{GAN} has not yet probed this intermediate class-space during its training and therefore we are reliant on the generator having learned any underlying class space relationships between the 5 training classes. The results show that for each case that the generated signals show distinct characteristics of the respective class pairs at most stages of the transition. We note that transitions in some cases appear to be rather abrupt, e.g., between the Gaussian pulse and the \ac{BBH}, \chris{but we point out that this feature, whilst not uncommon, is a strong function of the random latent-space location.}
%
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/generations/sg-rd.png}
    \includegraphics[width=\textwidth]{figures/generations/rd-wnb.png}
    \includegraphics[width=\textwidth]{figures/generations/wnb-blip.png}
    \includegraphics[width=\textwidth]{figures/generations/blip-bbh.png}
    \caption{\ac{GAN} generated class interpolated waveforms as a function of time showing class-space interpolations. A single latent space vector is used for all generations and is chosen randomly in the latent space. Each row shows generations using linearly interpolated classes as inputs to the generator. By row top to bottom: Sine-Gaussian to Ringdown, Ringdown to white noise burst, Gaussian pulse to BBH.\chris{add class vectors on each panel}}
    \label{fig:c_interp}
\end{figure}
%
\subsection{General points within the class space}
We have shown that the \acp{GAN} latent-space and class-space have structure that can be navigated via interpolation between pairs of locations within each respective space. Taking a step further, we can sample from the class-space in novel ways to create new inputs for the generator. These new points are categorised by the method used to sample from the class-space. The methods we use are divided into the following: 
%
\begin{itemize}
%
\item {\bf Vertex}: Points that lie at the corners of the five dimensional class space. These class-space locations are equivalent to the training set locations and are our closest generated representation of the training set.
%
\item {\bf Uniform}: Each of the entries in the class vector is sampled from a uniform distribution U[0,1]. This is equivalent to sampling uniformly within the 5-dimensional one-hit encoding hyper-cube.
%
\item {\bf Simplex}: This class vector we define as uniformly sampled points on a simplex, which is a generalization of a triangle in $k$-dimensions. We sample uniformly on the $k=4$ simplex that is embedded in the 5-dimensional class hyper cube. In practice we use the equivalent of sampling points from a $k=4$ Dirichlet distribution.
%
\end{itemize}

The vertex points are the most straightforward where one entry contains 1 and the other entries are zero and are equivalent to the class vectors that the GAN is trained on e.g. $\mathbf{c} = [1,0,0,0,0]$ would correspond to a sine-Gaussian generation. Uniform class vectors sampled from a uniform distribution and act like a random draw from the five dimensional hypercube. Uniformly sampling contains points up to a distance of unit from the closest class e.g. [0,0,0,0,0] is of distance 1 away from all classes (as is [1,1,1,1,1]). For simplex class vectors, we sample from the simplest hypersurface that intersects all the classes and has a symmetry such that no class is favoured over any other. For our 5-dimensional case this would correspond to a 4-simplex manifold. Sampling from the simplex can be seen as sampling from between the classes within the space.

In \cref{fig:simplexd_samples} we show generations conditioned on class vectors drawn from the 4-simplex. There are large variations in the signals with some looking like definitive classes, although this can be partially explained through the random draws from the simplex as there is finite probability that one class entry will dominate over the others. For instance the generations that look more like sine-Gaussians than unmodeled generally have a larger value placed in the first entry than others. Similarly \cref{fig:uniform_samples} shows generations conditioned on class vectors drawn uniformly in the unit hypercube. These types of generations tend to exhibit more noise and some are very low amplitude before being re-scaled. Artifacts like the signal not lining up to zero on the y axis is a result of re-scaling a particularly low amplitude signal. Both methods of generating unmodeled data, however, do show signals that look unrelated to the training data.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample1.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample2.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample3.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample4.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample5.png}
    \caption{GAN generations where the class vectors are sampled from the 4-D
plane intersecting all classes. Latent space locations for all signals are taken randomly from a Gaussian distribution and the signals are then re-scaled such that they peak at unity. The class label for each generation is shown above each pannel.}
    \label{fig:simplexd_samples}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample1.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample2.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample3.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample4.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample5.png}
    \caption{GAN generations where the class vectors are sampled uniformly in the hypercube class space. Latent space locations for all signals are taken randomly from a Gaussian distribution and the signals are then re-scaled such that they peak at unity.}
    \label{fig:uniform_samples}
\end{figure}

>>>>

% What this paper does on GANs in 1 paragraph - the point of the paper
%
In this work we aim to explore the use of \ac{ML} to generate and interpret
unmodeled \ac{GW} burst waveforms. Using the generative machine learning
model, \acp{GAN}, we train on five classes of waveforms in the time domain. Working on the assumption that \acp{GAN} construct smooth
 high dimensional vector spaces between their input and output, we can then
explore the space between the five classes to construct new
hybrid waveforms. As all the computationally expensive
processes occur during training, once trained, the model is able to
generate waveforms in fractions of a second and produce waveforms that are difficult to generate with current
techniques. These new varieties of waveforms can then be used to diagnose \michael{not sure about diagnose, maybe test or evaluate?}
detection algorithms, gain new insight into sources of \ac{GW}
bursts and  allow us to better train our algorithms on a
broader range of possible signals and therefore enhance our detection ability.

% the structure of the paper
%
This paper is organised as follows. In \cref{ML overview} we introduce the basic ideas of machine learning and discuss the choice of algorithm we used. In \cref{Method} we describe the training data and the details of the model. We present the results of the GAN in \cref{results} and show how unmodeled signals can be produced by interpolating and sampling within latent and class spaces. In \cref{cnn classifier} we show that a convolutional neural network \michael{should this have the accronym? Tried to fix it but it only puts the plural for me} classifier can be trained to distinguish between sets of our GAN generated waveforms from noise only cases. We conclude with a summary of the presented work in \cref{conclusions}.

@@

\ac{ML} algorithms aim to learn apparent relationships held within given data or `training
data' in order to make accurate predictions without the need for additional
programming. A common approach in \ac{ML} relies on the model learning from
past experience to make decisions on future events. Artificial neural networks are universal function approximators that are built from many single
processing units called neurons. The simplest neural network is the perceptron
layer~\cref{fig:perceptron} which shows a single neuron that takes a vector of real
inputs $x_{i},\ldots, x_{n}$
and maps them to an output according to the linear function, 
%
\begin{align}
f(x) = \sigma(\sum_i w_i x_i + b),
\label{eqn:neuron}
\end{align}
%
where $w$ and $b$ are the weights and bias and $\sigma$
denotes the activation function. The weights are numbers which can be thought
of as the strength between connected neurons. The output of a neuron is defined by its activation function which controls how the neuron `fires' depending on its input. Some examples of commonly used activation functions are shown in~\cref{fig:activations}. It is often useful to introduce a bias, $b$, such that the neuron remains inactive
above zero but is active when the sum reaches a defined threshold.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Generative Adversarial Networks}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% basic intro to GANs
%
A subset of deep learning that has seen fruitful development in recent years
are generative adversarial networks \acp{GAN}~\cite{Goodfellow2014}. These unsupervised algorithms learn patterns in a
given training data set using an adversarial process. The generations from
\acp{GAN} are currently state-of-the-art in fields such as high quality image
fidelity~\cite{brock2018large,karras2019analyzing}, text-to-image
translation~\cite{reed2016generative}, and video
prediction~\cite{liang2017dual} as well as time series
generations~\cite{esteban2017realvalued}.
%
% Basic components of a GAN
%
\acp{GAN} train two competing neural networks, consisting of a discriminator
network that is set up to distinguish between real and fake data and a
generator network that produces fake versions of the real data. The generator model performs a mapping from a fixed length vector $\mathbf{z}$ to its
representation of the data. The input vector is drawn randomly from a Gaussian distribution which is referred to as a latent space comprised of latent variables. The latent space is a compressed representation of a data distribution which the generator applies meaning to during training. Sampling points from this space allows the generator to produce a variety of different generations, with different points corresponding to different features in the generations. The discriminator maps its input $\mathbf{x}$ to a probability that the input came from either the training (real) data or
generator (fake).
\begin{comment}
Chris' definition of latent space if needed: It maps the intrinsic
variation of the training space onto the Gaussian latent space distribution.
\end{comment}
%
% training a GAN
%
During training, the discriminator and generator are updated using batches of data. Random latent vectors are given to the generator to produce a batch of fake samples and an equal batch of real samples is taken from the training data. The discriminator makes predictions on the real and fake samples and the model is updated through minimising the binary cross-entropy function \cite{Goodfellow-et-al-2016}
%
\begin{equation}
    L = -y \log(\hat{y}) - (1 - y) \log(1-\hat{y}),
    \label{eqn:crossentropy}
\end{equation}
where $\hat{y}$ is the network prediction and $y$ is the true output. While training the discriminator, D, on real data, we set $y = 1$ and $\hat{y} = D(\mathbf{x})$ which from \cref{eqn:crossentropy} gives $L(D(\mathbf{x}),1) = \log(D(\mathbf{x}))$. While training on fake data produced by the generator, G, $y = 0$ and $\hat{y} = D(G(\mathbf{z}))$ and so, $L(D(G(\mathbf{z})),0) = \log(1-(D(G(\mathbf{z}))))$. Since the objective of the discriminator is to correctly classify fake and real data these equations should be maximised, while the goal of the generator should be to minimize these equations. This gives the \ac{GAN} loss as
%
\begin{equation}
   \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{\text{data}}}(\mathbf{x})} [\text{log} D(\mathbf{x})] + \mathbb{E}_{\mathbf{z} \sim p_{\mathbf{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z})))],
 \label{equation:GANloss}
 \end{equation}
where $p_{\text{\text{data}}}(\mathbf{x})$ is the distribution of real data and $p_{\text{z}}(\mathbf{z})$ is the latent distribution.

\subsection{Training stages}
%
Training a GAN is involves updating both the discriminator and generator in stages. First, the discriminator is updated on real instances from the training set. We set the true label $y=1$ and calculate the loss with respect to the predictions $\hat{y}$ via \cref{equation:GANloss}. Stochastic gradient descent is used to minimize the loss which has reduced to $L_D(\textrm{real}) = -\log(\hat{y})$. The discriminator is then trained on fake instances taken from the generator where we set $y=0$ and minimize $L_D(\textrm{fake}) = -\log(1-\hat{y})$. To train the generator, we use a composite model of the generator and discriminator and allow the gradients to flow through this entire model. At this step we set the labels for the generator as real $y=1$ and minimize $L_G = -\log(\hat{y})$. The act of switching the labels and minimizing this form of loss is identical to the previously mentioned generator loss but is beneficial as in practice this function provides more useful gradients. This tweak to the generator loss is called non-saturating generator loss and was reported in the original \ac{GAN} paper \cite{Goodfellow2014}. It was also shown in that paper that if the generator and discriminator can no longer improve, then the discriminator can no longer distinguish between real and fake i.e. $D(x) = \frac{1}{2}$.

As \acp{GAN} are trained by updating one model at the expense of the other, they can be hard to train. GANs attempting to replicate complicated structures that do not
have the necessary architecture either struggle to produce results at all or
fall into the common failure mode know as mode collapse; where the generator
produces a small variety of samples or simply memorises the training set. The goal of \ac{GAN} training is to find an equilibrium between the two models, if this cannot be found then it is said that the \ac{GAN} has failed to converge. One way to diagnose \acp{GAN} during the development process is to keep track of the loss and accuracy over time. Loss plots, for example, as seen in \cref{fig:lossplot} can help to identify common failure modes or to check if the \ac{GAN} has indeed converged. Accuracy is another metric that may be used to monitor convergence and is defined as the number of correct predictions made divided by total number of predictions. There is currently no notion of early stopping in \acp{GAN}, instead, training is halted after convergence and by visually inspecting the generations. 
\begin{figure}[h!]
    \centering
    \includegraphics[width=0.85\textwidth]{figures/losses.png}
    \caption{Plot of the discriminator and generator loss and accuracy as a function of epochs. Early in training the losses oscillate as both models attempt to find their equilibrium, after which, both losses vary around a point which signifies stable training. Accuracies on the real and fake data are similar showing that neither model is stronger than the other.}
    \label{fig:lossplot}
\end{figure}

@@

% introduce CGANs and ACGANs
%
To gain more control over what a \ac{GAN} is able to generate, a conditional variant
of \acp{GAN} named \acp{CGAN}~\cite{cgan} was introduced by feeding in extra
information into the generator and discriminator such as a class label or
attribute label, $\mathbf{c}$. This simple addition has shown to work well in practice, for instance in image-to-image translation~\cite{isola2016imagetoimage}. We defined the classes by one-hot encoding, that is, each class resides at the corner points of a 5 dimensional cube. For example $\mathbf{c}=[0,1,0,0,0]$ represents the ringdown signal class. The training data and labels are drawn from a joint distribution $p_{\text{data}}(\mathbf{x},\mathbf{c})$, whereas when generating fake data we sample from $\mathbf{c}$ and $p_{\mathbf{z}}(\mathbf{z})$ independently. \cref{equation:GANloss} is modified to include the class labels 
~
\begin{equation}
   \mathop{\text{min}}_{G}  \mathop{\text{max}}_{D} V(D,G) = \mathbb{E}_{\mathbf{x} \sim p_{\text{\text{data}}}(\mathbf{x})} [\text{log} D(\mathbf{x|c})] + \mathbb{E}_{\mathbf{z} \sim p_{\text{z}}(\mathbf{z})} [\text{log}(1-D(G(\mathbf{z|c})))].
 \label{equation:CGANloss}
 \end{equation}
Fig. \ref{fig:gan_comparison} shows the differences in inputs and outputs of a GAN compared with a \ac{CGAN}. We will be using a conditional GAN for this study.

@@

% introduce the training data
%
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/training-sample.png}
    \caption{Examples of the five different waveforms that were used in training the \ac{GAN} for this study. Values of the parameters were selected randomly from uniform distributions from \cref{Tab:training_parms}.}
    \label{fig:training_waveforms}
\end{figure}
% introduce the signal models
%
We propose a signal generation scheme using a \ac{CGAN} trained on burst-like
waveforms. We call this \texttt{McGANn}\footnote{https://github.com/jmcginn/McGANn} and it is a \ac{CGAN} trained on five signal classes each spanning a range
of prior signal parameters. The signal classes are:

@@

The design of the networks is influenced by \cite{Radford2015} in which they use a \ac{DCGAN} architecture. The generator model is fully convolutional,
upsampled using strided transposed convolutions
with batch normalisation in the
first layer and ReLU activations throughout with the exception of a linear activation for
the output layer. The use of a linear activation guarantees the output can have negative and positive outputs. Each transposed
convolutional layer uses a kernel size of 18 and stride of 2. The discriminator network
mirrors that of the generator without batch normalization, using
LeakyReLU activations, SpatialDropout, and a 2-stride convolution for
downsampling. The discriminator output is a single node with sigmoid activation that can be interpreted as a probability of the the signal being real and both models are trained with binary cross
entropy \cref{eqn:crossentropy}. The full architecture description can be seen in~\cref{Tab:gan_training_parms}.

All models were designed with the Python Keras library \cite{} and TensorFlow \cite{} and trained on a GeForce RTX 2080 Ti GPU. We train the networks for 500 epochs which takes $\mathcal{O}(10)$ hours and save the model at each epoch. We choose an appropriate model by visually inspecting the generations at a point of convergence on the loss plot. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results} \label{results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

While the \ac{GAN} is trained on distinct one-hot encoded classes, we may test arbitrary points in the 5-dimensional class space to produce indistinct or hybrid waveforms. In order to explore the class-space, in \cref{fig:c_interp} we show results where the latent vector is held constant but we instead linearly
interpolate within the one-hot encoded class-space between pairs of the well-defined training class locations. In this scenario we highlight that the \ac{GAN} has not yet probed this intermediate class-space during its training and therefore we are reliant on the generator having learned any underlying class space relationships between the 5 training classes. The results show that for each case that the generated signals show distinct characteristics of the respective class pairs at most stages of the transition. We note that transitions in some cases appear to be rather abrupt, e.g., between the Gaussian pulse and the \ac{BBH}, \chris{but we point out that this feature, whilst not uncommon, is a strong function of the random latent-space location.}
%
\begin{figure}[!h]
    \centering
    \includegraphics[width=\textwidth]{figures/generations/sg-rd.png}
    \includegraphics[width=\textwidth]{figures/generations/rd-wnb.png}
    \includegraphics[width=\textwidth]{figures/generations/wnb-blip.png}
    \includegraphics[width=\textwidth]{figures/generations/blip-bbh.png}
    \caption{\ac{GAN} generated class interpolated waveforms as a function of time showing class-space interpolations. A single latent space vector is used for all generations and is chosen randomly in the latent space. Each row shows generations using linearly interpolated classes as inputs to the generator. By row top to bottom: Sine-Gaussian to Ringdown, Ringdown to white noise burst, Gaussian pulse to BBH.\chris{add class vectors on each panel}}
    \label{fig:c_interp}
\end{figure}
%
\subsection{General points within the class space}
We have shown that the \acp{GAN} latent-space and class-space have structure that can be navigated via interpolation between pairs of locations within each respective space. Taking a step further, we can sample from the class-space in novel ways to create new inputs for the generator. These new points are categorised by the method used to sample from the class-space. The methods we use are divided into the following: 
%
\begin{itemize}
%
\item {\bf Vertex}: Points that lie at the corners of the five dimensional class space. These class-space locations are equivalent to the training set locations and are our closest generated representation of the training set.
%
\item {\bf Simplex}: This class vector we define as uniformly sampled points on a simplex, which is a generalization of a triangle in $k$-dimensions. We sample uniformly on the $k=4$ simplex that is embedded in the 5-dimensional class hyper-cube. In practice we use the equivalent of sampling points from a $k=4$ Dirichlet distribution. It is useful to think of the simplex as the hyper-plane that intersects all 5 training classes. It is a subspace of the Uniform method.  
%
\item {\bf Uniform}: Each of the entries in the class vector is sampled from a uniform distribution $\text{U}[0,1]$. This is equivalent to sampling uniformly within the 5-dimensional one-hot encoding hyper-cube.
%
\end{itemize}

The vertex points are the most straightforward where one element of the class vector contains 1 and the other elements are zero. these points are equivalent to the class vectors that the \ac{GAN} is trained on e.g., $\mathbf{c} = [1,0,0,0,0]$ would correspond to a sine-Gaussian generation. Uniform class vectors with each element sampled from a uniform distribution are equivalent to a random draw from a 5-dimensional hyper-cube. Uniformly sampling generates class-space locations up to a maximum distance of unity from the closest class e.g. [0,0,0,0,0] is of distance 1 away from all classes (as is [1,1,1,1,1]). For simplex class vectors, we sample from the simplest hyper-surface that intersects all the classes and has a symmetry such that no training class location (any vertex) is favoured over any other. For our 5-dimensional case this corresponds to a 4-simplex manifold. Sampling from the simplex can be seen as sampling from the simplest space that spans between the training classes.

In \cref{fig:simplexd_samples} we show generations conditioned on class vectors drawn randomly from the 4-simplex. There are large variations in the signals with some having characteristics strongly resembling the training classes, although this can be partially explained through the random draws from the simplex as there is finite probability that one class entry will dominate over the others (i.e., the class-space location is close to a vertex). For instance the generations that look more like sine-Gaussians than hybrid waveforms generally have a larger value placed in the first class-space element than others. Similarly \cref{fig:uniform_samples} shows generations conditioned on class vectors drawn uniformly in the unit hyper-cube. These types of generations tend to exhibit more noise and some tend to be generated with very low amplitude prior to being re-scaled to have maximum amplitude of unity. Both methods of generating hybrid waveforms, however, do produce signals that appear to share characteristics from the training set but still be distinct in signal morphology. Upon inspection of a larger collection of waveform generations from both methods we do see a tendency for the uniform hyper-cube approach to generate more wider variety of hybrid waveforms that are more visually distinct from the training set. This is to be expected given that the simplex class-space is a subset of the hyper-cube and does not explore regions of the class-space as far from the training set vertices.

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample1.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample2.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample3.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample4.png}
    \includegraphics[width=\textwidth]{figures/generations/simplex_sample5.png}
    \caption{GAN generations where the class vectors are sampled from the 4-dimensional plane (simplex) intersecting all training classes. Latent space locations for all signals are taken randomly from a 100-dimensional Gaussian distribution and the signals are then re-scaled such that they have maximum absolute amplitude at unity. The class label for each generation is shown above each panel.}
    \label{fig:simplexd_samples}
\end{figure}

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample1.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample2.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample3.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample4.png}
    \includegraphics[width=\textwidth]{figures/generations/uniform_sample5.png}
    \caption{GAN generations where the class vectors are sampled uniformly in the hyper-cube class space. Latent space locations for all signals are taken randomly from a 100-dimensional Gaussian distribution and the signals are then re-scaled such that they have maximum absolute amplitude at unity.}
    \label{fig:uniform_samples}
\end{figure}

\section{CNN burst classifier} \label{cnn classifier}
In this section we develop a basic search analysis using a \ac{CNN} in order to compare the sensitivity of such a search to different \ac{GAN} generated waveforms in additive noise. We train a \ac{CNN} to perform simple classification and to distinguish between two classes: signals in additive Gaussian noise and Gaussian noise only. We are primarily interested in the relative sensitivity as a function of the types of waveforms used for training the network. We are also interested in how these differently trained networks perform when applied to data from waveform generations not used in the training process.
